---
title: "Modelling Data"
author: "Salim Damerdji"
output: rmarkdown::github_document
---

```{r}
 iat_df <- read.csv('../data/clean/iat_scores.csv') %>%
    as_tibble()
```

# Power Analysis

Specify a minimum difference of 1%
```{r}
quantile(iat_df$diff, c(.5,.55))
var(iat_df$diff)
```

```{r}
get_power <- function(n){
  df1 <- 3
  df2 <- 4*n-6
  ncp <- (n*.15^2)/(2*.10^2)
  cutoff <- qf(p=.95,df1=df1, df2=df2)
  return(pf(q=cutoff,df1=df1, df2=df2, ncp=ncp, lower.tail=FALSE))
}

powers <- sapply(6:30, get_power)
ggplot()+
  geom_line(aes(x=6:30, y=powers))+
  xlab('# Observations per Subgroup')+
  ylab('Power')+
  ggtitle('Power Analysis for Hypothesis that Any Two Treatment Groups Differ by D') +
  geom_vline(xintercept=12, color = 'dark green') +
  annotate('text',14.5,.84,label='n=12, power=.85')
```
# Exploratory Analysis

## Checking Greenwald's recommendation that we divide IAT score by pooled sd

```{r}
# People with more variable reaction times tended  to have larger differences between stage 3 and 5
cor(abs(iat_df$diff), iat_df$sd)
```
This confirms Greenwald's claim that there is a strong relationship between sd of reaction times and the absolute value of the difference in mean reaction times for stage 3 and 5.

## Plot reaction times from stages 1 through 5
```{r}
# As we expect, reaction times jump up when a new stage begins.
ggplot(no_outliers, aes(trial,response, group=time)) +
  geom_point(alpha=.05, size=.01)+
  geom_line(alpha=.1, size=.1)+
  xlab("Trial")+ 
  ylab("Response time in seconds") +
  ggtitle("After removing outliers: Each subject's response time during the experiment")+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


# Blocking
```{r}
# Blocking by gender was a good call
ggplot(iat_df) +
  geom_point(aes(x=factor, y = iat_d, color = gender))
iat_df
```



## Contrast: Treatments vs Control

Let's examine the contrast for whether there was a difference in response between the control group and the treatment groups. This corresponds to the coefficient vector of (-3, -3, 1, 1, 1, 1, 1, 1). In that vector, the first two entries correspond to the two control groups (one per gender); the remaining four entries correspond to the remaining treatment / block combinations.

```{r}
iat_df %>%
  group_by(factor, gender) %>%
  summarise(mean(diff))
```

Thus, the unbiased estimate is $3*(.0421) -3*(.2727) + .1218 + .1884 - .2159 + .2474 + .1271 + .2801$, which equals 0.0571. We use MS residuals as an estimate for error, by multiplying it by:

```{r}
3*(.0421) -3*(.2727) + .1218 + .1884 - .2159 + .2474 + .1271 + .2801
(-3)^2/6 + (-3)^2/6 + 1/6 + 1/6 + 1/6 + 1/6 + 1/6
```


Our t-statistic then is: 0.0571/sqrt(.1727 * 3.833333)

```{r}
0.0571/sqrt(.1691 * 3.833333)
```

```{r}
(1-pt(.07, df = 48-6))*2
```

```{r}
summary(aov(iat_d ~ prodiversity + antistereotype + prodiversity:antistereotype + gender, data = iat_df))

```

## ANOVA Assumptions:


## Residual Check

```{r fit_resid}
# Fitted vs. Residual
mod = aov(iat_d ~ prodiversity + antistereotype + gender, data = iat_df)

fit_resid = tibble(f = fitted(mod), r = residuals(mod))

fit_resid

iat_df <- mutate(iat_df, factor = as.factor(factor))

levels(iat_df$factor) <- c('Neither', 'Pro-diversity','Anti-stereotype','Both')
ggplot(iat_df, aes(x=fit_resid$f, y=fit_resid$r)) + 
  geom_jitter(width = 0.25,aes(color = gender, shape = factor)) + 
  geom_hline(yintercept = 0, color='red') + 
  ggtitle("Jittered Residuals against Fitted Values with Groups Labelled") + 
  xlab("Fitted Values") + 
  ylab("Residuals")+
  labs(subtitle = "Figure 5")


model.tables(mod)
```

### Independence
```{r}
# Is response time determined by when they took the IAT
ggplot(iat_df, aes(x=time, y=fit_resid$r)) + 
  geom_jitter(width = 0.25,aes(color = gender, shape = factor)) + 
  ylab("Residuals") +
  xlab("Time of Experiment") +
  ggtitle('Did Time of Trial Affect Residual')+
  scale_x_continuous(breaks = c(.5,.625, .75, .875), 
                     labels = c('12 PM', '3 PM', '6 PM','9 PM'))
```

```{r}
# Is response time determined by the computer on which they took the IAT
ggplot(iat_df, aes(x=cp, y=fit_resid$r)) + 
  geom_jitter(width = 0.25,aes(color = gender, shape = factor)) + 
  ylab("Residuals") +
  xlab("Computer") +
  ggtitle('Did CP of Trial Affect Residual')
```

Neither assumption is violated.